humanCdGF_predict_cfu

### 2016-10-12 code/HumanCdGF_predict_cfu.Rmd
colonization rate prediction

### Reframing the question for predicting CFU
#### 1. Can Day 0 community predict the median steady-state-colonized CFU?
#### 2. Can Day 0 community predict the rate of colonization?

@@

>>>>

### Reframing the question for predicting CFU
#### 1. Can Day 0 community predict the median steady-state-colonized CFU?

@@

#### 2. Can Day 0 community predict the rate of colonization?

```{r predict colonization rate, fig.width=10,fig.height=10}

meta_data %>% 
  filter(!is.na(day), !cage_id == 'NP1', 
         Early_Euth == FALSE, cdiff_strain == 431) %>% 
  ggplot(aes(x = factor(day), y = log_cfu, color = cage_id)) + 
    geom_line(aes(group = mouse_id), size=.25, alpha=0.4) + geom_point() +
    stat_summary(fun.y=mean, geom="line", aes(group = cage_id)) + 
    theme_bw() + labs(x = 'Day', y = 'C Diff CFU (log10)') +
    theme(legend.justification=c(1,0), legend.position=c(1,0), 
          legend.title=element_blank())

mouse_list <- meta_data %>% 
  filter(!is.na(day), !cage_id == 'NP1', 
         Early_Euth == FALSE, cdiff_strain == 431) %>% 
  select(mouse_id) %>% 
  unique()
mouse_list <- as.character(mouse_list$mouse_id)

col_rate <- c()
for(i in mouse_list){
  test <- meta_data %>% 
    filter(!is.na(day), mouse_id == i) %>% 
    select(day, log_cfu) %>% 
    filter(!is.na(log_cfu))
  if(1 %in% test$day){
    colonization_ratio <- test[test$day==1, 'log_cfu']/median(head(sort(test$log_cfu, decreasing = T),3))
  } else {
    colonization_ratio <- median(head(test[order(test$day), "log_cfu"],2))/
                            median(head(sort(test$log_cfu, decreasing = T),3))
  }
  col_rate <- rbind(col_rate, c(mouse_id = i,  ratio = colonization_ratio))
}

col_rate <- data.frame(mouse_id = col_rate[ , 1], ratio = as.numeric(col_rate[ , 2]))

col_rate <- meta_data %>% 
  select(mouse_id, cage_id) %>% 
  unique() %>% 
  right_join(col_rate)

shared_col_rate.df <- col_rate %>% 
    right_join(community_d0.df) %>% 
    select(-one_of(cage_source_features))
shared_colonized.df <- select(shared_col_rate.df, -mouse_id, -cage_id)

set.seed(seed)
rf_col_rate <- randomForest(ratio ~ ., data = shared_colonized.df,
                           importance = TRUE, ntree = n_trees, mtry = 14)

importance_sorted <- sort(importance(rf_col_rate)[,1], decreasing = T)
top_important_OTU <- rownames(data.frame(importance_sorted[importance_sorted/max(importance_sorted) > 0.6]))

top_col_rate_df <- data.frame(shared_colonized.df %>% 
  select(ratio, one_of(top_important_OTU)))

set.seed(seed)
rf_col_rate_features <- randomForest(ratio ~ ., data=top_col_rate_df,
                              importance=TRUE, ntree=n_trees, mtry = 2,
                              keep.forest=TRUE)

importance_final <- sort(importance(rf_col_rate_features)[,1], decreasing = T)

rf_col_rate_prediction <- data.frame(observed = shared_colonized.df$ratio, 
                       predicted = rf_col_rate_features$predicted)
rf_cr_prediction <- cbind(cage_id = shared_col_rate.df[,'cage_id'], rf_col_rate_prediction)

ggplot(data = rf_cr_prediction, aes(x=observed, y=predicted, color = cage_id)) + geom_point() + 
  labs(x= 'Observed', y = 'Predicted',
       title = "Random Forest Prediction of Relative Colonization level of C Difficile from Day 0 Community") + 
  theme_bw() + xlim(0,1) + ylim(0,1) + 
  annotate('text', x=0.15*(max(rf_cr_prediction$observed)), 
           y=0.95*(max(rf_cr_prediction$predicted)), 
           label=paste0('R^2 = ', round(tail(rf_col_rate_features$rsq, 1), 2)))

cage_color <- rainbow(length(unique(rf_cr_prediction$cage_id)))
for (i in seq_along(importance_final)) {
  partial_otu <- 
    partialPlot(rf_col_rate_features, top_col_rate_df, names(importance_final)[i],
                xlab='', main=paste(tax_label[i]), 
                ylim = c(min(top_col_rate_df$ratio),max(top_col_rate_df$ratio)))
  for(m in seq_along(unique(rf_cr_prediction$cage_id))){
    cage <- unique(rf_cr_prediction$cage_id)[m]
    points(y = top_col_rate_df[rf_cr_prediction$cage_id %in% cage, 'ratio'],
         x = top_col_rate_df[rf_cr_prediction$cage_id %in% cage , names(importance_final)[i]], 
         pch = 21, bg = cage_color[m])
  }
}
